---
title: "Markdown Assignment 17"
author: "John Jenkins"
date: "Fall 2024"
output: html_document
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(leaflet)
library(dygraphs)
library(xts)
library(DT)
library(scatterplot3d)
library(threejs)
groundhogs<-read_csv("groundhogs.csv")
library(nycflights13)

```

### **Question 1**
```{r}
cities<-tibble(names=c("Seattle","Houston","Orlando","Boston","Liberty","New York City"), lat=c(47.608013,29.749907,28.538336,42.361145,39.246096, 40.730610), long=c(-122.335167,-95.358421,-81.379234,-71.057083, -94.418469, -73.935242))
cities<- cities %>% mutate(stopnumber=str_c("Stop #",row_number(),": ",names))


leaflet(cities) %>% addTiles() %>% addMarkers(~long,~lat, label=~names, popup=~stopnumber)
```

***
### **Question 2**
```{r}
groundhogpic<-makeIcon("groundhogpic.png","groundhogpic.png",40,40)

groundhogs

leaflet(groundhogs) %>% addTiles() %>% addMarkers(~longitude,~latitude, label=~name, icon=groundhogpic, clusterOptions=markerClusterOptions())
```

> A fun name of one of the ground hogs is Octoraro Orphie.

***
### **Question 3**
```{r}
newground<-groundhogs %>% select(-(description))
datatable(newground,editable=list(target="row",disable=list(columns=3:4)))
```

***
### **Question 4**
```{r}
newfl<-flights %>% group_by(time_hour) %>% summarize(averagedist=mean(distance,na.rm=TRUE))

flightsTimeSeries <- xts(x=newfl$averagedist,order.by=newfl$time_hour)

dygraph(flightsTimeSeries) %>% dySeries("V1",label="Average Distance")
```

> One oddity I notice in the plot is that around July the higher average distace traveled is more condensed. One pattern I notice is a high concentration of average flights being between 800 and 1200 miles.

***
### **Question 5**
```{r}
newdiamond<-diamonds %>% filter(z<8)

scatterplot3js(newdiamond$x,newdiamond$y,newdiamond$z, color=as.numeric(newdiamond$cut),size=0.35)
```

> It can be concluded that neither of the x, y, or z variables have a relationship to the cut variable.